---
title: "qm3"
output: pdf_document
date: "2023-11-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ageing Population vs Pension and Healthcare Costs

# Project Overview

This R Markdown document presents an analysis of the relationship between the ageing population, pension expenditure, and healthcare costs in France. The analysis is conducted using a variety of datasets related to population demographics, economic indicators, pension funds, and healthcare expenditures. The primary goal of this project is to explore the impact of an ageing population on pension expenditure and healthcare costs and to identify any potential correlations or trends that may exist.

## Data Sources

The project uses several datasets, including:

- **Population and Age Demographics:**
  - `population_growth_rate.csv`: Population growth rates in France.
  - `birth_rates.csv`: Birth rates in France.
  - `life_expectancy.csv`: Life expectancy data.
  - `age-dependency.csv`: Age dependency ratios.

- **Economic Data:**
  - `unemployment-rate.csv`: Unemployment rates.
  - `france_gdp.csv`: Gross Domestic Product (GDP) data for France.

- **Pensions:**
  - `pension_funds.csv`: Data related to pension funds and expenditure.

- **Healthcare:**
  - `total-healthcare-expenditure-gdp.csv`: Total healthcare expenditure as a percentage of GDP.
  - `france-healthcare-spending.csv`: Healthcare spending data in France.

## Data Analysis

The analysis involves the following steps:

1. Data Import: The project begins with the import of the necessary datasets, ensuring data consistency and cleaning if required.

2. Exploratory Data Analysis (EDA): Initial data exploration is conducted to gain insights into the distributions and characteristics of the datasets.

3. Correlation Analysis: The project examines the correlations between various demographic and economic factors, such as population growth, birth rates, and unemployment, and their impact on pension and healthcare costs.

4. Multiple Linear Regression: A multiple linear regression model is used to predict pension and healthcare costs based on various independent variables.

5. Visualization: Data visualizations are created to illustrate key findings and relationships.

6. Conclusion: The project concludes with a summary of findings and insights regarding the relationship between ageing population, pension expenditure, and healthcare costs in France.

# Importing the data
```{r}
data_folder <- "~/Downloads/"

# Define the list of CSV files
csv_files <- c("population_growth_rate.csv", "birth_rates.csv", "life_expectancy.csv","age-dependency.csv","unemployment-rate.csv", "france_gdp.csv", "pension_funds.csv", "total-healthcare-expenditure-gdp.csv", "france-healthcare-spending.csv","age_distribution.csv"
 
)

# Population and Age Demographics

#Fine as is (1950-2100)
population_growth_rate <- read.csv(file.path(data_folder, "population_growth_rate.csv"))

#Per 1000 people (1950-2100)
birth_rates <- read.csv(file.path(data_folder, "birth_rates.csv"))

#Fine to use as is (1950-2100)
life_expectancy <- read.csv(file.path(data_folder, "life_expectancy.csv"))

# Work with the data as is, just extract data from France (1950-2021)
age_dependency <- read.csv(file.path(data_folder, "age-dependency.csv"))

# Percentages (2006-2022)
age_distribution <- read.csv(file.path(data_folder, "age_distribution.csv"))


# Economic Data
# Use France Data (1991-2021)
unemployment_rate <- read.csv(file.path(data_folder, "unemployment-rate.csv"))

# Billion US dollars (1987-2028)
gdp <- read.csv(file.path(data_folder, "france_gdp.csv"))


# Pensions
#Percentages of GDP Share (2002-2021)
pension_funds <- read.csv(file.path(data_folder, "pension_funds.csv"))

# Health 
#Percentages of GDP share, use only France's Data (2000-2019)
healthcare <- read.csv(file.path(data_folder, "total-healthcare-expenditure-gdp.csv"))

#Percentage of GDP (2000-2020)
france_healthcare <- read.csv(file.path(data_folder, "france-healthcare-spending.csv"))


```

## Cleaning, Formalising Data and Creating DataFrames

You can also embed plots, for example:

```{r pressure, echo=FALSE}

age_dependency_data <- subset(age_dependency, Entity == "France", select = c("Year", "Child.dependency.ratio...Sex..all...Age..none...Variant..estimates"))
head(age_dependency_data)

```
```{r}
# Assuming 'birth_rates' is your data frame
birth_rates$date <- as.numeric(format(as.Date(birth_rates$date, format = "%m/%d/%y"), "%Y"))

# Reset the row names (index)
rownames(birth_rates) <- NULL

# Drop the 'date' column
birth_rates <- birth_rates[-which(names(birth_rates) == 'date')]

# Create a new 'year' column from 1950 to 2100
birth_rates$year <- 1950:2100


head(birth_rates)

```
```{r}
france_healthcare$date <- as.numeric(format(as.Date(france_healthcare$date, format = "%m/%d/%y"), "%Y"))

# Reset the row names (index)
rownames(france_healthcare) <- NULL


head(france_healthcare)

```
```{r}
france_healthcare <- france_healthcare[-which(names(france_healthcare) == 'X')]
head(france_healthcare)
```
```{r}
# Assuming 'birth_rates' is your data frame
life_expectancy$date <- as.numeric(format(as.Date(life_expectancy$date, format = "%m/%d/%y"), "%Y"))

# Reset the row names (index)
rownames(life_expectancy) <- NULL

# Drop the 'date' column
life_expectancy <- life_expectancy[-which(names(life_expectancy) == 'date')]

# Create a new 'year' column from 1950 to 2100
life_expectancy$year <- 1950:2100


head(life_expectancy)
```
```{r}
pension_funds <- pension_funds[-which(names(pension_funds) == 'X')]
head(pension_funds)
```
```{r}
# Assuming 'birth_rates' is your data frame
population_growth_rate$date <- as.numeric(format(as.Date(population_growth_rate$date, format = "%m/%d/%y"), "%Y"))

# Reset the row names (index)
rownames(population_growth_rate) <- NULL

# Drop the 'date' column
population_growth_rate <- population_growth_rate[-which(names(population_growth_rate) == 'date')]

# Create a new 'year' column from 1950 to 2100
population_growth_rate$year <- 1950:2100


head(population_growth_rate)

```
```{r}
unemployment_rate_data <- subset(unemployment_rate, Entity == "France", select = c("Year", "Unemployment..total....of.total.labor.force...modeled.ILO.estimate."))
head(unemployment_rate_data)

```
## Summary statistics for new dataframes
```{r}
# Summary statistics for population_growth_rate
summary(population_growth_rate)

# Summary statistics for birth_rates
summary(birth_rates)

# Summary statistics for age_distribution
summary(age_distribution)

# Summary statistics for life_expectancy
summary(life_expectancy)

# Summary statistics for age_dependency
summary(age_dependency_data)

# Summary statistics for unemployment_rate
summary(unemployment_rate_data)

# Summary statistics for gdp
summary(gdp)

# Summary statistics for pension_funds
summary(pension_funds)

# Summary statistics for france_healthcare
summary(france_healthcare)

```

# Multiple Linear Regression:
Dependent Variable: Pension Expenditure or Healthcare Costs.
Independent Variables: Various indicators of an aging population (e.g., percentage of elderly population, dependency ratio, life expectancy) and other relevant predictors (e.g., GDP, unemployment rate).
This model allows you to assess the impact of multiple predictors on pension expenditure and healthcare costs simultaneously.

```{r}
population_growth_rate <- population_growth_rate %>%
  filter(year >= 2006, year <= 2019)
birth_rates <- birth_rates %>%
  filter(year >= 2006, year <= 2019)
life_expectancy<- life_expectancy %>%
  filter(year >= 2006, year <= 2019)
age_dependency <- age_dependency %>%
  filter(Year >= 2006, Year <= 2019)
age_distribution <- age_distribution %>%
  filter(year >= 2006, year <= 2019)
unemployment_rate_data <- unemployment_rate_data %>%
  filter(Year >= 2006, Year <= 2019)
gdp <- gdp %>%
  filter(year >= 2006, year <= 2019)
pension_funds <- pension_funds %>%
  filter(year >= 2006, year <= 2019)
france_healthcare <- france_healthcare %>%
  filter(date >= 2006, date <= 2019)
# Formalize the year column
population_growth_rate <- population_growth_rate %>%
  mutate(year = as.integer(year))

birth_rates <- birth_rates %>%
  mutate(year = as.integer(year))

life_expectancy <- life_expectancy %>%
  mutate(year = as.integer(year))

age_dependency_data <- age_dependency_data %>%
  mutate(year = as.integer(Year))  # Rename Year to year

age_distribution <- age_distribution %>%
  mutate(year = as.integer(year))

unemployment_rate_data <- unemployment_rate_data %>%
  mutate(year = as.integer(Year))  # Rename Year to year

gdp <- gdp %>%
  mutate(year = as.integer(year))

pension_funds <- pension_funds %>%
  mutate(year = as.integer(year))

france_healthcare <- france_healthcare %>%
  mutate(year = as.integer(date))  # Rename date to year


```


```{r}
# Print the head of each dataset
cat("Head of population_growth_rate dataset:\n")
print(head(population_growth_rate))

cat("\nHead of birth_rates dataset:\n")
print(head(birth_rates))

cat("\nHead of life_expectancy dataset:\n")
print(head(life_expectancy))

cat("\nHead of age_dependency dataset:\n")
print(head(age_dependency))

cat("\nHead of age_distribution dataset:\n")
print(head(age_distribution))

cat("\nHead of unemployment_rate dataset:\n")
print(head(unemployment_rate_data))

cat("\nHead of gdp dataset:\n")
print(head(gdp))

cat("\nHead of pension_funds dataset:\n")
print(head(pension_funds))

cat("\nHead of france_healthcare dataset:\n")
print(head(france_healthcare))

```
```{r}
# Print the dimensions of each dataframe
cat("Dimensions of population_growth_rate dataset:\n")
print(dim(population_growth_rate))

cat("\nDimensions of birth_rates dataset:\n")
print(dim(birth_rates))

cat("\nDimensions of life_expectancy dataset:\n")
print(dim(life_expectancy))

cat("\nDimensions of age_dependency dataset:\n")
print(dim(age_dependency_data))

cat("\nDimensions of age_distribution dataset:\n")
print(dim(age_distribution))

cat("\nDimensions of unemployment_rate dataset:\n")
print(dim(unemployment_rate_data))

cat("\nDimensions of gdp dataset:\n")
print(dim(gdp))

cat("\nDimensions of pension_funds dataset:\n")
print(dim(pension_funds))

cat("\nDimensions of france_healthcare dataset:\n")
print(dim(france_healthcare))

```



```{r}
library(dplyr)

resulting_df <- population_growth_rate %>%
  left_join(birth_rates, by = "year") %>%
  left_join(life_expectancy, by = "year") %>%
  left_join(age_dependency_data, by = "year") %>%
  left_join(age_distribution, by = "year") %>%
  left_join(unemployment_rate_data, by = "year") %>%
  left_join(gdp, by = "year") %>%
  left_join(pension_funds, by = "year") %>%
  left_join(france_healthcare, by = "year")

# View the first few rows of the resulting dataframe
head(resulting_df)
```
```{r}
# Rename the variables in the resulting dataframe more concisely
resulting_df <- resulting_df %>%
  rename(
    Population = Population,
    GrowthRate = Annual.Growth.Rate,
    Year = year,
    BirthsPer1000 = Births.per.1000.People,
    AnnualChange_x = Annual...Change.x,
    LifeExpectancy = Life.Expectancy.from.Birth..Years.,
    AnnualChange_y = Annual...Change.y,
    Year_x = Year.x,
    ChildDependency = Child.dependency.ratio...Sex..all...Age..none...Variant..estimates,
    Age0_19 = X0.to.19.years.old,
    Age20_59 = X20.to.59.years.old,
    Age60_64 = X60.to.64.years.old,
    Age65_older = X65.years.and.older,
    Year_y = Year.y,
    Unemployment = Unemployment..total....of.total.labor.force...modeled.ILO.estimate.,
    GDP = France..Gross.domestic.product..GDP..in.current.prices.from.1987.to.2028..in.billion.U.S..dollars.,
    PensionShareofGDP = pension.funds.as.a.GDP.share,
    Date = date,
    PerCapitaUS = Per.Capita..US...,
    GDPShare = X..of.GDP
  )

# View the first few rows of the resulting dataframe with renamed variables
head(resulting_df)

```

```{r}
# Fit a multiple linear regression model for healthcare costs
healthcare_model <- lm(GDPShare ~ Age65_older + ChildDependency + Unemployment, data =resulting_df)

# Summary of the healthcare model
summary(healthcare_model)
```

```{r}


# Residuals vs. Fitted Values Plot
plot(healthcare_model)
# Normal Q-Q Plot
qqnorm(residuals(healthcare_model))
qqline(residuals(healthcare_model))
# Scale-Location Plot
plot(fitted(healthcare_model), sqrt(abs(resid(healthcare_model))))
# Residuals vs. Leverage Plot
plot(hatvalues(healthcare_model), resid(healthcare_model))

# Partial Regression Plots
library(car)  # Load the 'car' package for the 'qqPlot' function
qqPlot(healthcare_model, "Age65_older")
qqPlot(healthcare_model, "ChildDependency")
qqPlot(healthcare_model, "Unemployment")



```

```{r}
# Create a linear regression model with only "PensionShareGDP" and "Age65AndOlder"
linear_model <- lm(GDPShare ~ Age65_older, data = resulting_df)

# Summarize the regression results
summary(pension_age_model)
# Quadratic Regression
quadratic_model <- lm(GDPShare ~ Age65_older + I(Age65_older^2), data = resulting_df)

# Logarithmic Regression
logarithmic_model <- lm(log(GDPShare) ~ log(Age65_older), data = resulting_df)

# Evaluate the models
linear_r_squared <- summary(linear_model)$r.squared
quadratic_r_squared <- summary(quadratic_model)$r.squared
logarithmic_r_squared <- summary(logarithmic_model)$r.squared

# Print R-squared values
cat("Linear R-squared:", linear_r_squared, "\n")
cat("Quadratic R-squared:", quadratic_r_squared, "\n")
cat("Logarithmic R-squared:", logarithmic_r_squared, "\n")

# Choose the model with the highest R-squared
best_model <- linear_model
if (quadratic_r_squared > linear_r_squared) {
  best_model <- quadratic_model
}
if (logarithmic_r_squared > max(linear_r_squared, quadratic_r_squared)) {
  best_model <- logarithmic_model
}


# Visualize the results
plot(resulting_df$Age65_older, resulting_df$GDPShare, xlab = "Age 65 and Older", ylab = "Healthcare Share of GDP", main = "Regression Models")

# Fit and plot the linear regression line
abline(linear_model, col = "blue")

# Fit and plot the quadratic regression curve
quadratic_fit <- predict(quadratic_model)
lines(resulting_df$Age65_older, quadratic_fit, col = "red")

# Fit and plot the logarithmic regression curve
logarithmic_fit <- exp(predict(logarithmic_model))
lines(resulting_df$Age65_older, logarithmic_fit, col = "green")

# Add legends
legend("topleft", legend = paste("Linear R-squared =", round(summary(linear_model)$r.squared, 4)), col = "blue", lty = 1)
legend("topleft", legend = paste("Quadratic R-squared =", round(summary(quadratic_model)$r.squared, 4)), col = "red", lty = 1)
legend("topleft", legend = paste("Logarithmic R-squared =", round(summary(logarithmic_model)$r.squared, 4)), col = "green", lty = 1)

```


```{r}
# Linear Regression
linear_model <- lm(PensionShareofGDP ~ Age65_older, data = resulting_df)

# Quadratic Regression
quadratic_model <- lm(PensionShareofGDP ~ Age65_older + I(Age65_older^2), data = resulting_df)

# Logarithmic Regression
logarithmic_model <- lm(log(PensionShareofGDP) ~ log(Age65_older), data = resulting_df)

# Evaluate the models
linear_r_squared <- summary(linear_model)$r.squared
quadratic_r_squared <- summary(quadratic_model)$r.squared
logarithmic_r_squared <- summary(logarithmic_model)$r.squared

# Print R-squared values
cat("Linear R-squared:", linear_r_squared, "\n")
cat("Quadratic R-squared:", quadratic_r_squared, "\n")
cat("Logarithmic R-squared:", logarithmic_r_squared, "\n")

# Choose the model with the highest R-squared
best_model <- linear_model
if (quadratic_r_squared > linear_r_squared) {
  best_model <- quadratic_model
}
if (logarithmic_r_squared > max(linear_r_squared, quadratic_r_squared)) {
  best_model <- logarithmic_model
}



# Visualize the results
plot(resulting_df$Age65_older, resulting_df$PensionShareofGDP, xlab = "Age 65 and Older", ylab = "Pension Share of GDP", main = "Regression Models")

# Fit and plot the linear regression line
abline(linear_model, col = "blue")

# Fit and plot the quadratic regression curve
quadratic_fit <- predict(quadratic_model)
lines(resulting_df$Age65_older, quadratic_fit, col = "red")

# Fit and plot the logarithmic regression curve
logarithmic_fit <- exp(predict(logarithmic_model))
lines(resulting_df$Age65_older, logarithmic_fit, col = "green")

# Add legends
legend("topleft", legend = paste("Linear R-squared =", round(summary(linear_model)$r.squared, 4)), col = "blue", lty = 1)
legend("topleft", legend = paste("Quadratic R-squared =", round(summary(quadratic_model)$r.squared, 4)), col = "red", lty = 1)
legend("topleft", legend = paste("Logarithmic R-squared =", round(summary(logarithmic_model)$r.squared, 4)), col = "green", lty = 1)

```


```{r}
# Fit a multiple linear regression model with more variables
linear_model <- lm(PensionShareofGDP ~ Age65_older + ChildDependency + GDPShare + Unemployment, data = resulting_df)

# Summary of the updated linear model
summary(linear_model)


# Residuals vs. Fitted Values Plot
plot(linear_model)
# Normal Q-Q Plot
qqnorm(residuals(linear_model))
qqline(residuals(linear_model))
# Scale-Location Plot
plot(fitted(linear_model), sqrt(abs(resid(linear_model))))

# Partial Regression Plots
library(car)  # Load the 'car' package for the 'qqPlot' function
qqPlot(linear_model, "Age65_older")
qqPlot(linear_model, "ChildDependency")
qqPlot(linear_model, "Unemployment")


```

```{r}
# Fit a multiple linear regression model with more variables
linear_model <- lm(PensionShareofGDP ~ Age65_older + ChildDependency + GDPShare + Unemployment+ Population + BirthsPer1000, data = resulting_df)

# Summary of the updated linear model
summary(linear_model)

```

```{r}
# Fit a multiple linear regression model with more variables
linear_model <- lm(GDPShare ~ Age65_older + ChildDependency + GDPShare + Unemployment+ Population , data = resulting_df)

# Summary of the updated linear model
summary(linear_model)
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
